---
title: "biopixR - Introduction"
author: "Tim Brauckhoff"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{biopixR - Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      cache = FALSE, comment = NA, verbose = TRUE, 
                      fig.width = 5, fig.height = 5, dev = 'jpeg', 
                      dev.args = list(quality = 25))
options(digits = 3)
```

# Getting started

The biopixR package includes an image of micro particles (beads) as an example to demonstrate its analytical and processing abilities for biological imagery. This sample image displays the package's features, enabling users to experiment with image analysis and manipulation within the contexts of biotechnology and life sciences. Researchers and practitioners can utilize this illustration to comprehend the application of biopixR to their individual imaging requirements, whether pertaining to cell biology, microscopy, or any other biological imaging applications.

## detecting objects in an image

```{r beads}
library(biopixR)
plot(beads)
```

As the biopixR package primarily uses the image processing packages imager and magick, the examples are stored in the specific imager class: cimg.

```{r class}
class(beads)
```

The objective of this task is to extract important information from an image consisting of beads. As a preliminary step, it is essential to distinguish between individual beads and acquire their corresponding coordinates or positions. The `objectDetection` function uses edge detection and labeling techniques to gather elaborate information about the beads, allowing for identification and differentiation of individual particles according to their unique edges. This procedure assists in deriving precise coordinates for every bead in the image, which sets the groundwork for subsequent analysis and characterization of the micro particles within the biopixR package.

```{r objectDetection}
res_objectDetection <- objectDetection(beads, alpha = 0.75, sigma = 0.1)
```

This function generates a list of objects. Let's examine the specific outcomes and explore methods for visualizing them, starting with the center coordinates of the beads:

```{r visualization_1}
plot(beads)
with(res_objectDetection$centers, 
     points(res_objectDetection$centers$mxx,
            res_objectDetection$centers$myy,
            col = factor(res_objectDetection$centers$value),
            pch = 19))
```

Upon observation, it is apparent that each individual bead is correctly identified with a singular point at its center and differentiated by varying colors to indicate their distinctiveness, which meets our intended objective. Nevertheless, markings of clotted beads, which are later referred to as doublets or multiplets, do not adhere to the expected pattern. Doublets being identified as a single entity is a logical consequence of their edges vanishing along the contact surface, resulting in the observed behavior. The multiplets make this matter more complex. The challenge in accurately identifying the edges between beads is similar to that of doublets. On the other hand, the `objectDetection` function utilizes labeling post edge detection, employing a filling process similar to the EBImage `fillHull` function. In this step, the function does not recognize the space within the multiplet as a single area to be filled, resulting in an unfilled space. Then, during the subsequent labeling, multiple 'centers' are assigned to the still-present edges, generating several detected points along the edges of the multiplet. The next visualization displays the latter phenomenon.

Let's examine the next output from `objectDetection`. This function captures the coordinates of labeled regions, providing precise details about the position of each bead. By leveraging another function within the package, `changePixelColor`, we can selectively color specific coordinates in a cimg. Thus, we can apply this function to tint all the extracted coordinates in the bead image and assess whether the outcome aligns with our expectations.

```{r visualization_2}
changePixelColor(beads, res_objectDetection$coordinates, color = "purple")
```

As expected, all individual beads (singlets) and the doublet are fully colored, indicating the successful collection of all related coordinates. As previously mentioned, the collection procedure solely evaluates the contours of the multiplet. Although unintentional, this result offers an advantage as it is crucial to exclude multiplets and doublets from subsequent analysis due to their close proximity. As they are in close proximity, there is a possibility of inducing fluorescence in one another, thereby producing misleading signals and contributing to false positives during analysis.

The function incorporates an internal visualization feature that utilizes the `drawcircles` function from the imager package to outline circles around the identified centers, replicating the initial plot. 

```{r visualization_3}
res_objectDetection$marked_beads |> plot()
```

Nonetheless, a crucial differentiation occurs in obtaining the highlighted beads as a cimg, which opens up possibilities for the creation of an interactive tool using tcltk. This step facilitates the development of an interactive interface, empowering users to dynamically explore the adjustment of various variables and observe the corresponding shifts in detected centers. The interactive interface is presented through the `interactive_objectDetection` function within the biopixR package.

## filter beads according to size and proximity

As stated previously, it is crucial to remove doublets and multiplets prior to analysis. This objective will be addressed in the next section, where we will use the `proximityFilter` to achieve this goal. The function works by examining each collected center and scanning a specific radius for additional centers. If another center is detected within this range, both are discarded due to their close proximity. This method is especially efficient in dealing with multiplets as the identified centers along the contours are frequently found in close proximity. Furthermore, even beads that were at first identified as singles are tested to ensure that there are no neighboring centers within the defined range. Users can customize the radius for this examination as required.

```{r}
res_proximityFilter <- proximityFilter(res_objectDetection, radius = 10)
```

**visualization proximityFilter**:

```{r}
plot(beads)
with(res_proximityFilter$remaining.centers, 
     points(res_proximityFilter$remaining.centers$mxx,
            res_proximityFilter$remaining.centers$myy,
            col = "darkgreen",
            pch = 19))
with(res_proximityFilter$discard,
     points(res_proximityFilter$discard$mx,
            res_proximityFilter$discard$my,
            col = "darkred",
            pch = 19))
```

As expected, the centers that were found along the contours of the multiplet are now eliminated due to their proximity. To conduct a small experiment, I will increase the radius for the subsequent plot, intending to discard not only the multiplet but also the two beads in the upper-middle section.

```{r}
small_experiment <- proximityFilter(res_objectDetection, radius = 13)

plot(beads)
with(small_experiment$remaining.centers, 
     points(small_experiment$remaining.centers$mxx,
            small_experiment$remaining.centers$myy,
            col = "darkgreen",
            pch = 19))
with(small_experiment$discard,
     points(small_experiment$discard$mx,
            small_experiment$discard$my,
            col = "darkred",
            pch = 19))
```

After the successful completion of the small experiment, our attention now returns to the main objective of removing the doublets. To achieve this, the `sizeFilter` will be employed on the remaining beads, utilizing both the centers and all coordinates. The `sizeFilter` algorithm calculates the number of pixels in each labeled object, allowing the user to set individual limits to discard doublets. In scenarios where the contour of the multiplet might be perceived as a single region, the removal based on distance alone may not be effective. Therefore, the `sizeFilter` provides an additional option by allowing the establishment of a lower limit.

```{r}
res_sizeFilter <- sizeFilter(res_proximityFilter,
                             lowerlimit = 0,
                             upperlimit = 150)
```

**visualization sizeFilter**:

```{r}
changePixelColor(beads, res_sizeFilter$remaining.coordinates.s, color = "darkgreen")
```

The objective of this chapter has been accomplished, as demonstrated by the most recent graph. After eliminating the multiplet with the `proximityFilter`, the doublet has now been removed from the dataset using the `sizeFilter`. To present this outcome, the `changePixelColor` function was once again employed, coloring every remaining pixel. Subsequently, the remaining beads are highlighted in dark green, indicating their successful passage through the filtering process.

## displaying results

To conclude this chapter, we need to extract meaningful information from the filtered data set. One of the most fundamental results to be displayed after applying a filter is undoubtedly the number of remaining and discarded objects. As the size of the objects has already been analyzed through the `sizeFilter`, this information should also be included in the display. Moreover, the intensity of the signal is a crucial parameter for both beads and any fluorescent image. Finally, it may be of interest to calculate the area density, which represents the number of beads relative to the pixel area of the entire image. To extract this information, the `ResultAnalytics` function from the biopixR package is utilized. This function requires the data frame of the remaining coordinates, the cluster size, and the original image obtained from the `sizeFilter` function as inputs.

```{r}
result <- ResultAnalytics(res_sizeFilter)
result$detailed
```

While it's possible to showcase a detailed version of results, featuring individual beads with their cluster number, size, intensity, and coordinates, this presentation method can become quite overwhelming, especially when dealing with larger images containing numerous objects. Consequently, the image results are summarized in a single row, emphasizing the key parameters described earlier.

```{r}
result$Summary
```

A final note regarding the estimated number of rejected objects: A careful observer may have noticed that the actual count of discarded beads is 7. Explaining this discrepancy requires investigating the algorithm, which is prone to an inevitable deviation. Since the beads are not easily distinguished through straightforward labeling, simple counting, which would offer the most accurate result, is not feasible. Instead, the algorithm uses an automatic `threshold` to collect all TRUE coordinates and counts them. This count is then divided by the mean size of the beads, and after subtracting the count of passing beads, it yields the number of discarded objects. The potential issue arises from overlapping. This is because when beads overlap, their shared area is counted only once in the image, leading to a lower discarded bead count. It is crucial to consider this when relying on this parameter.

### relative distance

Another important aspect to consider in bioimaging is the relative distance between individual objects. By examining the spatial relationships between the identified objects, researchers can gain valuable insights into their organization and distribution within the image. This information is critical for understanding biological phenomena such as spatial patterns, cell clustering, or the arrangement of microstructures. In the following sections, we present a method that explores the euclidean distance between objects in order to enable users to investigate and quantify spatial relationships within their bioimages.

To begin this task, we first summarize the remaining coordinates after applying the `sizeFilters` using the data.table package. This process involves computing the centers, which is also a part of the `objectDetection` algorithm. The calculated centers then serve as the initial reference points for measuring the distances between objects.

```{r}
DT <- data.table(res_sizeFilter$remaining.coordinates.s)
res_center <- DT[, .(x = mean(x), y = mean(y)), by = cluster]
```

The relative distance between multiple objects typically refers to the distances between each pair of objects in relation to each other. The method to calculate relative distance depends on the dimensionality of your objects (e.g., 1D, 2D, 3D) and the specific metric you want to use (e.g., Euclidean distance, Manhattan distance). Here, I'll provide a general approach for 2D objects using the Euclidean distance. Assuming you have a set of points in a 2D space (e.g., x, y coordinates), you can calculate the euclidean distance between each pair of points. The formula for euclidean distance between two points \((x_1, y_1)\) and \((x_2, y_2)\) is given by:

\[ \text{Distance} = \sqrt{(x_2 - x_1)^2 + (y_2 - y_1)^2} \]

```{r}
euclidean_distance <- function(point1, point2) {
  sqrt((point2$x - point1$x)^2 + (point2$y - point1$y)^2)
}
```

Now, let's apply this function on our generated centers. Initially, we create an empty matrix to store the results. Next, we develop a nested for loop to systematically compare each center with every other center, with the exception of comparing a center to itself, resulting in a missing value. This calculation results in a distance matrix.

```{r}
num_points <- nrow(res_center)
relative_distances <- matrix(NA, nrow = num_points, ncol = num_points)

for (i in 1:num_points) {
  for (j in 1:num_points) {
    if(i == j) {
      next
      } else {
        relative_distances[i, j] <- euclidean_distance(res_center[i, ], res_center[j, ])
    }
  }
}

as.matrix(relative_distances)
```

After removing missing values, the average relative distance for each center is calculated, which provides individual insight into its distance from the other beads. The resulting values are then added as a new column to the previously presented detailed results, providing further information about the distinct properties of each bead.

```{r}
for (a in 1:num_points) {
  result$detailed$relative_distance[a] <- mean(na.omit(relative_distances[a, ]))
}

result$detailed
```

As mentioned previously, the detailed table results for a larger image would unavoidably become overly extensive. The same issue is applicable to the matrix. Therefore, to overcome this issue, instead of relying solely on the detailed table results, the mean relative distance is calculated and included in the summary table.

```{r}
result$Summary$mean_distance <- mean(result$detailed$relative_distance)
result$Summary
```

# Case study: beads in droplets - *in progress*

## dealing with discontinous edges

## characterizing partitions 
