---
title: "biopixR - Introduction"
author: "Tim Brauckhoff"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{biopixR - Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE,
  cache = FALSE, comment = NA, verbose = TRUE,
  fig.width = 5, fig.height = 5, dev = "jpeg",
  dev.args = list(quality = 25)
)
options(digits = 3)
```

# Getting started

The biopixR package includes an image of microbeads as an example to demonstrate
its analytical and processing abilities for biological imagery. This sample
image displays the package's features, enabling users to experiment with image
analysis and manipulation within the contexts of biotechnology and life
sciences. Researchers and practitioners can utilize this illustration to
comprehend the application of biopixR to their individual imaging requirements,
whether pertaining to cell biology, microscopy, or any other biological imaging
applications.

## Detecting objects in an image

```{r beads}
library(biopixR)
plot(beads)
```

As the biopixR package primarily uses the image processing packages imager and
magick, the examples are stored in the specific imager class: cimg.

```{r class}
class(beads)
```

The objective of this task is to extract important information from an image
consisting of microbeads. As a preliminary step, it is essential to distinguish
between individual microbeads and acquire their corresponding coordinates or
positions. The `objectDetection` function uses edge detection and labeling
techniques to gather elaborate information about the microbeads, allowing for
the identification and differentiation of individual particles according to
their unique edges. This procedure assists in deriving precise coordinates for
every microbead in the image, which sets the groundwork for subsequent analysis
and characterization of the microbeads within the biopixR package.

```{r objectDetection}
res_objectDetection <- objectDetection(beads, alpha = 0.75, sigma = 0.1)
```

This function generates a list of objects. Let's examine the specific outcomes
and explore methods for visualizing them, starting with the center coordinates
of the microbeads:

```{r visualization_1}
plot(beads)
with(
  res_objectDetection$centers,
  points(res_objectDetection$centers$mxx,
    res_objectDetection$centers$myy,
    col = factor(res_objectDetection$centers$value),
    pch = 19
  )
)
```

Upon observation, it is apparent that each individual microbead is correctly
identified with a singular point at its center and differentiated by varying
colors to indicate their distinctiveness, which meets our intended objective.
Nevertheless, the markings of clotted microbeads, which are later referred to as
doublets or multiplets, do not adhere to the expected pattern. Doublets being
identified as a single entity is a logical consequence of their edges vanishing
along the contact surface, resulting in the observed behavior. The multiplets
make this matter more complex. The challenge of accurately identifying the edges
between microbeads is similar to that of doublets. On the other hand, the
`objectDetection` function utilizes labeling post edge detection, employing a
filling process similar to the EBImage `fillHull` function. In this step, the
function does not recognize the space within the multiplet as a single area to
be filled, resulting in an unfilled space. Then, during the subsequent labeling,
multiple 'centers' are assigned to the still-present edges, generating several
detected points along the edges of the multiplet. The next visualization
displays the latter phenomenon.

Let's examine the next output from `objectDetection`. This function captures the
coordinates of labeled regions, providing precise details about the position of
each microbead. By leveraging another function within the package,
`changePixelColor`, we can selectively color-specific coordinates in a cimg.
Thus, we can apply this function to tint all the extracted coordinates in the
microbead image and assess whether the outcome aligns with our expectations.

```{r visualization_2}
changePixelColor(beads, res_objectDetection$coordinates, color = "purple")
```

As expected, all individual microbeads (singlets) and the doublet are fully
colored, indicating the successful collection of all related coordinates. As
previously mentioned, the collection procedure solely evaluates the contours of
the multiplet. Although unintentional, this result offers an advantage, as it is
crucial to exclude multiplets and doublets from subsequent analysis due to their
proximity. As they are in close proximity, there is a possibility of inducing
fluorescence in one another, thereby producing misleading signals and
contributing to false positives during analysis.

The function incorporates an internal visualization feature that utilizes the
`drawcircles` function from the imager package to outline circles around the
identified centers, replicating the initial plot.

```{r visualization_3}
res_objectDetection$marked_beads |> plot()
```

Nonetheless, a crucial differentiation occurs in obtaining the highlighted
microbeads as a cimg, which opens up possibilities for the creation of an
interactive tool using tcltk. This step facilitates the development of an
interactive interface, empowering users to dynamically explore the adjustment of
various variables and observe the corresponding shifts in detected centers. The
interactive interface is presented through the `interactive_objectDetection`
function within the biopixR package.

## Filter microbeads according to size and proximity

As stated previously, it is crucial to remove doublets and multiplets prior to
analysis. This objective will be addressed in the next section, where we will
use the `proximityFilter` to achieve this goal. The function works by examining
each collected center and scanning a specific radius for additional centers. If
another center is detected within this range, both are discarded due to their
proximity. This method is especially efficient in dealing with multiplets as the
identified centers along the contours are frequently found in proximity.
Furthermore, even microbeads that were at first identified as singles are tested
to ensure that there are no neighboring centers within the defined range. Users
can customize the radius for this examination as required.

```{r}
res_proximityFilter <- proximityFilter(res_objectDetection, radius = 10)
```

**visualization proximityFilter**:

```{r}
plot(beads)
with(
  res_proximityFilter$remaining.centers,
  points(res_proximityFilter$remaining.centers$mxx,
    res_proximityFilter$remaining.centers$myy,
    col = "darkgreen",
    pch = 19
  )
)
with(
  res_proximityFilter$discard,
  points(res_proximityFilter$discard$mx,
    res_proximityFilter$discard$my,
    col = "darkred",
    pch = 19
  )
)
```

As expected, the centers that were found along the contours of the multiplet are
now eliminated due to their proximity. To conduct a small experiment, I will
increase the radius for the subsequent plot, intending to discard not only the
multiplet but also the two microbeads in the upper-middle section.

```{r}
small_experiment <- proximityFilter(res_objectDetection, radius = 13)

plot(beads)
with(
  small_experiment$remaining.centers,
  points(small_experiment$remaining.centers$mxx,
    small_experiment$remaining.centers$myy,
    col = "darkgreen",
    pch = 19
  )
)
with(
  small_experiment$discard,
  points(small_experiment$discard$mx,
    small_experiment$discard$my,
    col = "darkred",
    pch = 19
  )
)
```

After the successful completion of the small experiment, our attention now
returns to the main objective of removing the doublets. To achieve this, the
`sizeFilter` will be employed on the remaining microbeads, utilizing both the
centers and all coordinates. The `sizeFilter` algorithm calculates the number of
pixels in each labeled object, allowing the user to set individual limits to
discard doublets. In scenarios where the contour of the multiplet might be
perceived as a single region, removal based on distance alone may not be
effective. Therefore, the `sizeFilter` provides an additional option by allowing
the establishment of a lower limit.

```{r}
res_sizeFilter <- sizeFilter(res_proximityFilter,
  lowerlimit = 0,
  upperlimit = 150
)
```

**visualization sizeFilter**:

```{r}
DT <- data.table(res_sizeFilter$remaining.coordinates.s)
res_center <- DT[, .(x = mean(x), y = mean(y)), by = cluster]

changePixelColor(beads,
                 res_sizeFilter$remaining.coordinates.s,
                 color = "darkgreen")
text(res_center$x,
     res_center$y,
     res_center$cluster,
     col = "grey")
```

The objective of this chapter has been accomplished, as demonstrated by the most
recent graph. After eliminating the multiplet with the `proximityFilter`, the
doublet has now been removed from the dataset using the `sizeFilter`. To present
this outcome, the `changePixelColor` function was once again employed, coloring
every remaining pixel. Subsequently, the remaining microbeads are highlighted in
dark green, indicating their successful passage through the filtering process.

## Displaying results

To conclude this chapter, we need to extract meaningful information from the
filtered data set. One of the most fundamental results to be displayed after
applying a filter is undoubtedly the number of remaining and discarded objects.
As the size of the objects has already been analyzed through the `sizeFilter`,
this information should also be included in the display. Moreover, the intensity
of the signal is a crucial parameter for both microbeads and any fluorescent
image. Finally, it may be of interest to calculate the area density, which
represents the number of microbeads relative to the pixel area of the entire
image. To extract this information, the `resultAnalytics` function from the
biopixR package is utilized. This function requires the data frame of the
remaining coordinates, the cluster size, and the original image obtained from
the `sizeFilter` function as inputs.

```{r}
result <- resultAnalytics(res_sizeFilter)
result$detailed
```

While it's possible to showcase a detailed version of results featuring
individual microbeads with their cluster number, size, intensity, and
coordinates, this presentation method can become quite overwhelming, especially
when dealing with larger images containing numerous objects. Consequently, the
image results are summarized in a single row, emphasizing the key parameters
described earlier.

```{r}
result$summary
```

A final note regarding the estimated number of rejected objects: A careful
observer may have noticed that the actual count of discarded microbeads is 7.
Explaining this discrepancy requires investigating the algorithm, which is prone
to an inevitable deviation. Since the microbeads are not easily distinguished
through straightforward labeling, simple counting, which would offer the most
accurate result, is not feasible. Instead, the algorithm uses an automatic
`threshold` to collect all TRUE coordinates and count them. This count is then
divided by the mean size of the microbeads, and after subtracting the count of
passing microbeads, it yields the number of discarded objects. The potential
issue arises from overlapping. This is because when microbeads overlap, their
shared area is counted only once in the image, leading to a lower discarded
microbead count. It is crucial to consider this when relying on this parameter.

### Relative distance between microparticles

Another important aspect to consider in bioimaging is the relative distance
between individual objects. By examining the spatial relationships between the
identified objects, researchers can gain valuable insights into their
organization and distribution within the image. This information is critical for
understanding biological phenomena such as spatial patterns, cell clustering, or
the arrangement of microstructures. In the following sections, we present a
method that explores the Euclidean distance between objects to enable users to
investigate and quantify spatial relationships within their bioimages.

To begin this task, we first summarize the remaining coordinates after applying
the `sizeFilters` using the data.table package. This process involves computing
the centers, which is also a part of the `objectDetection` algorithm. The
calculated centers then serve as the initial reference points for measuring the
distances between objects.

```{r}
DT <- data.table(res_sizeFilter$remaining.coordinates.s)
res_center <- DT[, .(x = mean(x), y = mean(y)), by = cluster]
```

The relative distance between multiple objects typically refers to the distances
between each pair of objects in relation to each other. The method to calculate
relative distance depends on the dimensionality of your objects
(e.g., 1D, 2D, 3D) and the specific metric you want to use (e.g., Euclidean
distance, Manhattan distance). Here, I'll provide a general approach for 2D
objects using the Euclidean distance. Assuming you have a set of points in a 2D
space (e.g., x, y coordinates), you can calculate the Euclidean distance between
each pair of points. The formula for Euclidean distance between two points
((x\_1, y\_1)) and ((x\_2, y\_2)) is:

\[ \text{Distance} = \sqrt{(x_2 - x_1)^2 + (y_2 - y_1)^2} \]

```{r}
euclidean_distance <- function(point1, point2) {
  sqrt((point2$x - point1$x)^2 + (point2$y - point1$y)^2)
}
```

Now, let's apply this function on our generated centers. Initially, we create an
empty matrix to store the results. Next, we develop a nested for loop to
systematically compare each center with every other center, except for comparing
a center to itself, resulting in a missing value. This calculation results in a
distance matrix.

```{r}
num_points <- nrow(res_center)
relative_distances <- matrix(NA, nrow = num_points, ncol = num_points)

for (i in 1:num_points) {
  for (j in 1:num_points) {
    if (i == j) {
      next
    } else {
      relative_distances[i, j] <- euclidean_distance(res_center[i, ], res_center[j, ])
    }
  }
}

as.matrix(relative_distances)
```

After removing missing values, the average relative distance for each center is
calculated, which provides individual insight into its distance from the other
microbeads. The resulting values are then added as a new column to the
previously presented detailed results, providing further information about the
distinct properties of each microbead.

```{r}
for (a in 1:num_points) {
  result$detailed$relative_distance[a] <- mean(na.omit(relative_distances[a, ]))
}

result$detailed
```

As mentioned previously, the detailed table results for a larger image would
unavoidably become overly extensive. The same issue is applicable to the matrix.
Therefore, to overcome this issue, instead of relying solely on the detailed
table results, the mean relative distance is calculated and included in the
summary table. 

```{r}
result$summary$mean_distance <- mean(result$detailed$relative_distance)
result$summary
```

*will be available in the `resultAnalytics` function as of version 0.1.1*

# Case study: microbeads in droplets

To further illustrate the package's capabilities, the following section presents
a case study mainly focused on addressing discontinuous edges in image analysis.
The study showcases the integration of crucial data from two images to determine
the quantities of droplets and microbeads. Additionally, the analysis aims to
investigate the frequency of events in which a single microbead joins a droplet,
as opposed to situations in which multiple microbeads are present in a single
droplet. The study employs an algorithm that focuses on filling gaps along
discontinuous edges. This is achieved through a combination of detecting line
ends and interpolating pixels. By using this comprehensive method, the study
provides valuable perspectives on the distribution between droplets and
microbeads in the provided image. These findings demonstrate the flexibility of
the package to handle complex image analysis scenarios.

The following images serve as test subjects for the upcoming study. The first
image displays a brightfield view showing droplets, some of which contain
microbeads. The second image on the left displays the fluorescent channel,
exhibiting only the microbeads.

```{r fig.show='hold', out.width="49%"}
plot(droplets)
plot(droplet_beads)
```

## Dealing with discontinous edges

In typical fashion for image analysis, this study starts by applying a threshold
to the bright-field image. Subsequently, the resulting image uncovers a distinct
challenge: the edges of individual partitions are not continuous. In order to
differentiate individual partitions and evaluate whether they contain
microbeads, it is crucial to bridge these gaps. Fortunately, the package offers
a specialized function, `fillLineGaps`, to address this issue in image analysis.
This algorithm identifies line endpoints and connects them to the nearest
neighboring edge that is not their own. Additionally, the `objectDetection`
function can eliminate specific objects, such as microbeads. This step is
crucial for avoiding the unwanted outcome of line ends becoming connected to
microbeads. The code chunks below, derived from the `fillLineGaps` function,
visually demonstrating the elimination of microbeads and the detection of line
ends by highlighting them with the `changePixelColor` function.

```{r fig.show='hold', out.width="49%"}
# preprocessing: threshold, negate and mirroring
thresh <- threshold(droplets, "13%")
thresh_cimg <- as.cimg(thresh)
thresh_magick <- cimg2magick(thresh_cimg)
neg_thresh <- image_negate(thresh_magick)
neg_thresh_cimg <- magick2cimg(neg_thresh)
neg_thresh_m <- mirror(neg_thresh_cimg, axis = "x")

# first remove microbeads from droplet image (important for the linking of
# discontinuous edges, as otherwise they may connect with the microbeads)
# removes objects to prevent reconnecting with labeled regions that
# are not lines/edges
beads_to_del <- droplet_beads
bead_coords <-
  objectDetection(beads_to_del, alpha = 1, sigma = 0.1)

# transform binary image to array to modify individual values
thresh_array <- as.array(neg_thresh_m)
for (i in 1:nrow(bead_coords$coordinates)) {
  thresh_array[bead_coords$coordinates[i, 1],
               bead_coords$coordinates[i, 2], 1, 1] <- 0
}
# removed microbeads from droplets and retransformation to cimg
thresh_clean_cimg <- as.cimg(thresh_array)

# displaying problem of discontinous edges
plot(thresh_cimg)
# displaying removed microbeads
plot(thresh_clean_cimg)
```

```{r fig.align='center'}
# same orientation for 'cimg' and 'magick-image'
thresh_clean_m <- mirror(thresh_clean_cimg, axis = "x")
thresh_clean_magick <- cimg2magick(thresh_clean_m)

# getting coordinates of all line ends
mo1_lineends <- image_morphology(thresh_clean_magick,
                                 "HitAndMiss", "LineEnds")

# transform extracted coordinates into data frame
lineends_cimg <- magick2cimg(mo1_lineends)

end_points <- which(lineends_cimg == TRUE, arr.ind = TRUE)
end_points_df <- as.data.frame(end_points)
colnames(end_points_df) <- c("x", "y", "dim3", "dim4")

# highlighted line ends
vis_lineend <-
  changePixelColor(thresh_clean_cimg, end_points_df, color = "green")
```

After reviewing part of the `fillLineGaps` function, let us apply it to the
example images provided. The first three parameters have already been discussed,
which entail converting to binary using thresholding and eliminating identified
objects. The alpha and sigma parameters, denoting the threshold and smoothing
factor, are derived from the `cannyEdge` function in the imager package. Moving
on to the next parameter, the radius determines the maximum pixel range around
each line end that ought to be scanned for another edge. The iterations
parameter specifies the number of times the algorithm will be applied to the
given image. The function incorporates an internal visualization that highlights
the pixels added by the algorithm to fill the line gaps. In the following images
the visualization and the result of the function are displayed.

```{r fig.show='hold', out.width="49%"}
closed_gaps <- fillLineGaps(droplets,
  droplet_beads,
  threshold = "13%",
  alpha = 1,
  sigma = 0.1,
  radius = 5,
  iterations = 3,
  visualize = TRUE
)

closed_gaps |> plot()
```

As intended, the contours show reduced fragmentation and improved continuity,
enabling further analysis to extract meaningful information from the image.
While this accomplishment is remarkable, it is important to acknowledge the
algorithm's limitations. The algorithm performs admirably in situations where
relatively straight lines are fragmented, but challenges arise in more
complicated situations. One issue arises from diagonal line endings where, for
lines with a one-pixel width, each pixel is treated as a separate cluster. As a
result, the direct neighbor meets the reconnection requirements. To address this
problem, diagonal line endings will not reconnect with their own cluster or the
first direct neighbor's cluster. A different issue arises when there are
multiple edges in the scan area. In these instances, the endpoint will reconnect
with all of them, potentially generating new small partitions and a clotted-like
structure. Despite this, the algorithm successfully manages to close gaps in the
majority of cases, rendering it satisfactory for our specific example.

### Animated visualization of the `fillLineGaps` algorithm

```{r fig.align='center'}
first_img <- vis_lineend
second_img <- closed_gaps

first_m <- mirror(first_img, axis = "x")
second_m <- mirror(second_img, axis = "x")

first_magick <- cimg2magick(first_m)
second_magick <- cimg2magick(second_m)

img <- c(first_magick, second_magick)

image_animate(image_scale(img, "500x635"), fps = 1, dispose = "previous")
```


## Characterizing partitions

Having resolved the initial difficulty of the discontinuous edges, the focus
returns to the characterization of the partitions. Let's examine whether we can
now distinguish individual partitions after reducing the discontinuity. This is
achieved by labeling the 'closed_gaps' image. However, we should abstain from
visualizing the image at this point, as it would display a multitude of colors
since every pixel should be labeled.

```{r}
# label resulting image with filled gaps
lab_partitions <- label(closed_gaps)
df_lab_part <- as.data.frame(lab_partitions) |>
  subset(value > 0)
```

In the following section, the labeled edges will be discarded to isolate
individual partitions. This will be achieved by only taking into account
coordinates that are equal to zero. The end result will be displayed with the
use of the `changePixelColor` function, which colors every pixel except for the
contours. (It's worth noting that the top-left partition is excluded for unknown
reasons.)

```{r fig.align='center'}
# removing the edges as labeled regions
x <- list()
y <- list()
value <- list()

for (g in 1:nrow(df_lab_part)) {
  # droplets_array <- as.array(droplets)
  if (closed_gaps[df_lab_part$x[g], df_lab_part$y[g], 1, 1] == 0) {
    x[g] <- df_lab_part$x[g]
    y[g] <- df_lab_part$y[g]
    value[g] <- df_lab_part$value[g]
  }
}

# data frame of labeled partitions (without contours)
clean_lab_df <- data.frame(x = unlist(x),
                           y = unlist(y),
                           value = unlist(value))

changePixelColor(closed_gaps, clean_lab_df, color = "purple")
```

Let's summarize the clusters to validate their recognition as distinct
partitions. This is done by grouping the data by cluster and calculating the
mean coordinates using the data.table package. The resulting labeling of each
partition contains its cluster number in the center, and as desired, each
partition is distinguished by a unique cluster number.

```{r fig.align='center', out.width="100%"}
# summarizing data frame by cluster (value)
DT_droplet <- data.table(clean_lab_df)

grouped_droplets <-
  DT_droplet[, .(mx = mean(x), my = mean(y)), by = value]

plot(closed_gaps)
text(
  grouped_droplets$mx,
  grouped_droplets$my,
  grouped_droplets$value,
  col = "gray",
  cex = 0.5
)
```

Once the partitions have been identified, the main goal is to determine the
quantity of microbeads in each one. This can be achieved without analyzing every
single pixel by using the center coordinates of the microbeads. Each partition
is analyzed to confirm which contains coordinates that match a microbead center.
The result is shown in a brief table containing the cluster number in the first
row and the corresponding number of microbeads in each partition in the second
row.

```{r}
# checking in which partition the center coordinates of the microbeads are present
bead_partition <- list()
for (c in 1:nrow(bead_coords$centers)) {
  partition_pos <- which(
    clean_lab_df$x == round(bead_coords$centers$mxx[c]) &
      clean_lab_df$y == round(bead_coords$centers$myy[c])
  )
  bead_partition[c] <- clean_lab_df$value[partition_pos]
}
res_part <- table(as.character(bead_partition))
res_part
```

### Analysis of the partition size

This section analyzes partition sizes, which is critical due to the potential
creation of small inclusions by the `fillLineGaps` algorithm, leading to 'false'
partitions. The analysis involves counting the pixels within each partition and
generating a histogram, which reveals a positively skewed distribution
(`e1071::skewness(size_df$size)` = 3.63). As a result, the majority of
partitions appear to be small, with sizes less than 100 pixels.

```{r}
size_list <- list()
for (d in unique(clean_lab_df$value)) {
  pxl_number <- which(clean_lab_df$value == d)
  size <- length(pxl_number)
  size_list[d] <- size
}

size_df <- data.frame(partition = unique(clean_lab_df$value),
                      size = unlist(size_list))

hist(
  size_df$size,
  main = "Histogramm of partition size",
  xlab = "size",
  # labels = TRUE,
  breaks = "Scott",
  ylim = c(0, 120),
  xlim = c(0, 1400)
)
```

To reduce errors caused by incorrectly created partitions, any partitions with a
size of one pixel will be discarded. Although this may result in the removal of
some correctly reconnected small partitions, it is expected to improve accuracy
by eliminating certain falsely generated partitions.

```{r}
small <- which(size_df$size == 1)
number_small <- length(small)
```

The data frame will be adjusted in the next section by subtracting the count of
small partitions.

## Displaying result

As previously mentioned, the former table may be overly extensive and difficult
to extract information from at first glance. Therefore, a brief data frame is
produced, presenting only the most essential information collected from the two
images. Moreover, for reference, the initial image of droplets that contain the
microbeads is presented below.

```{r}
data.frame(
  partitions = nrow(grouped_droplets) - number_small,
  empty_partitions = nrow(grouped_droplets) - length(unique(bead_partition)) - number_small,
  bead_partitions = length(unique(bead_partition)),
  single_bead = length(which(res_part == 1)),
  muliple_beads = length(which(res_part > 1))
)
```

```{r fig.show='hold', out.width="49%"}
plot(droplets)
plot(closed_gaps)
```

