---
title: 'scanDir log file'
output: pdf_document
--- 
# Analysis of directory notebook/Images 
2024-05-13 17:54:40 Importing images from directory...   
2024-05-13 17:54:40 Checking md5sums...   
2024-05-13 17:54:40 Creating parallel backend...   
2024-05-13 17:54:41 Parallel backend successfully created. 6 workers recruited.  
2024-05-13 17:54:41 Starting image analysis...   
2024-05-13 17:54:41 Currently analyzing: notebook/Images/beads_large1.bmp   
2024-05-13 17:54:41 Currently analyzing: notebook/Images/beads.png   
2024-05-13 17:54:41 Currently analyzing: notebook/Images/beads_large2.png   
2024-05-13 17:54:41 Currently analyzing: notebook/Images/droplet_beads.png   
```{r, echo=FALSE}
directory_path <- getwd()
target_file_name <- 'result.RDS'
all_files <- list.files(directory_path, recursive = TRUE, full.names = TRUE)
matching_files <- all_files[grep(target_file_name, all_files, ignore.case = TRUE)]

md5result <- readRDS(matching_files)
name <- basename(md5result$file)
md5result <- data.frame(name = name,
                        amount = md5result$number_of_beads,
                        size = md5result$mean_size,
                        intensity = md5result$mean_intensity,
                        density = md5result$bead_density,
                        rejected = md5result$estimated_rejected,
                        coverage = md5result$coverage)
md5result
```  
```{r, echo=FALSE}
library(biopixR)
target_directory_name <- 'log_files'
all_directories <- list.dirs(directory_path, recursive = TRUE, full.names = TRUE)
matching_directories <- all_directories[grep(target_directory_name, all_directories, ignore.case = TRUE)]

for (i in 1:nrow(md5result)) {
file_path <- file.path(matching_directories, paste0('log_', i, '.RDS'))
import <- readRDS(file_path)
image <- import$image[[1]]
data <- import$data
name <- name[i]

plot(image, main = name)
with(data$unfiltered_centers, points(data$unfiltered_centers$mx, data$unfiltered_centers$my, col = 'darkred'))
with(data$detailed, points(data$detailed$x, data$detailed$y, col = 'darkgreen'))
}
```  
