---
title: 'scanDir log file'
output: pdf_document
--- 
# Analysis of directory notebook/Images/analysis 
2024-05-28 10:54:59 Importing images from directory...   
2024-05-28 10:55:00 Checking md5sums...   
2024-05-28 10:55:00 Currently analyzing: notebook/Images/analysis/1_000000_40x_Cy5_Grey=1000_Z=3509370_Well=A1_Ti=149999us_t=000173547ms.bmp   
2024-05-28 10:55:00 Starting object detection    
2024-05-28 10:55:01 Starting feature extraction    
2024-05-28 10:55:03 Analysis was successfully completed    
2024-05-28 10:55:03 Currently analyzing: notebook/Images/analysis/1_000007_40x_Cy5_Grey=1000_Z=3516170_Well=C1_Ti=149999us_t=000242390ms.bmp   
2024-05-28 10:55:04 Starting object detection    
2024-05-28 10:55:06 Starting feature extraction    
2024-05-28 10:55:09 Analysis was successfully completed    
2024-05-28 10:55:09 Currently analyzing: notebook/Images/analysis/1_000013_40x_Cy5_Grey=1000_Z=3537210_Well=E1_Ti=149999us_t=000315500ms.bmp   
2024-05-28 10:55:09 Starting object detection    
2024-05-28 10:55:12 Starting feature extraction    
2024-05-28 10:55:18 Analysis was successfully completed    
```{r, echo=FALSE}
directory_path <- getwd()
target_file_name <- 'result.RDS'
all_files <- list.files(directory_path, recursive = TRUE, full.names = TRUE)
matching_files <- all_files[grep(target_file_name, all_files, ignore.case = TRUE)]

md5result <- readRDS(matching_files)
name <- basename(md5result$file)
md5result <- data.frame(name = name,
                        amount = md5result$number_of_objects,
                        size = md5result$mean_size,
                        intensity = md5result$mean_intensity,
                        rejected = md5result$estimated_rejected,
                        coverage = md5result$coverage)
md5result
```  
```{r, echo=FALSE}
target_directory_name <- 'log_files'
all_directories <- list.dirs(directory_path, recursive = TRUE, full.names = TRUE)
matching_directories <- all_directories[grep(target_directory_name, all_directories, ignore.case = TRUE)]

file_names <- list.files(directory_path, full.names = TRUE)
file_names <- file_names[!file.info(file_names)$isdir]
file_names <-
  file_names[grep('\\.Rmd$', file_names, invert = TRUE)]
file_names <- tools::file_path_sans_ext(basename(file_names))

for (i in 1:nrow(md5result)) {
file_path <- file.path(matching_directories, paste0('log_', file_names[i], '.RDS'))
import <- readRDS(file_path)
image <- import$image[[1]]
data <- import$data
name_i <- name[i]

plot(image, main = name_i)
with(data$unfiltered_centers, points(data$unfiltered_centers$mx, data$unfiltered_centers$my, col = 'darkred'))
with(data$detailed, points(data$detailed$x, data$detailed$y, col = 'darkgreen'))
changePixelColor(image, data$coordinates, color = factor(data$coordinates$value), visualize = TRUE)
}

```  
